Epistemic Semantics in Guarded String Models

### 2020-06-17 tex/draft.tex
update algebraic stuff in section 2

Kleene Algebra with Tests has the algebraic signature
$\tuple{K,+,\cdot,*,\bar{},0,1}$.  To this we add an unary operation
$\Diamond_a$ for each agent, and a unary complent operation
$\cdot^{c}$ on propositions.  This results in the signature
$\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$ for
epistemic KAT with two agents.

We interpret each term in as a set of guarded strings as indicated in
Figure~\ref{fig:interp}. A bare event $e$ is interpreted as the set of
guarded strings allowed by $\varphi$ and $P$. The operation $p + q$ is
set union. The constant $0$ is the empty set.  The constant 1 is the
set of all truth assignments obeying the constraint on truth
assignments, i.e. $\{10,01\}$ in the coin example.  The operation
$\cdot$ is fusion product raised to sets: $x \cdot y$ is the set of
all defined fusion products of an element of $x$ with and element of
$y$.  The operation $*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$,
where $x^i$ is the $i$-times product of $x$ with itself and $x^0=1$.
Subsets of $1$ indicated by $t$ are also elements of $K$, and these
form the Boolean algebra of tests, denoting all atoms admitted by the
constraint $\varphi$ and the boolean test $t$.  We ``overload'' $\cdot$
to further indicate boolean conjunction, $+$ to further indicate
boolean disjunction, and $0,1$ to indicate the zero and one elements
of the boolean algebra. The overbar operation denotes complement in
this Boolean algebra: $\overline t$ denotes all atoms $\alpha$ such
that $\alpha \Rightarrow \neg t$.  Other boolean operators (such as
implication $\Rightarrow$) are encoded as syntactic sugar in the
standard way. The complement operation ${p}^c$ is complement at the
level of sets of guarded strings, defined to every set the universe
$\_^*$ except those denoted by $p$. Finally, the epistemic modality
$\Diamond_a$ is interpreted using Kripke semantics, as pre-image
relative to a fixed relation $R_a$ between guarded strings,
$\Diamond_a x = \setabs{u}{\exists v.\element{v}{x} \wedge u R_a v}$.
Here $u$ and $v$ are guarded strings, while $x$ is an element of $K$.

@@

It remains to define the Kripke relation on guarded strings from an
agent specification as in Figure \ref{epikfigure}. An an agent
specification pairs each bare event with a set of bare events, and so
determines a relation between bare events, call it relation $R_a$ for
an agent $a$.  This determines a relation $\hat{R_a}$ between
decorated events, see \ex{Rhat}.  It is generalized to a relation
$\doublehat{R}$ on arbitrary guarded strings by relational Kleene
star, as defined in \ex{Rhathat}. The operation $\cdot$ in the
definition is KAT product, which enforces matching of tests.  The
number of terms $n$ is construed as satisfying $n\geq 1$.  {\em What
  is desired for $n=0$?}

\begin{example}   \label{Rhat}
\hspace{-2mm} \(
\hat{R} =_{\mbox{\footnotesize{def.}}} 
\{\tuple{u,v} | \exists c \exists d. cRd \wedge \element{u}{\hat{c}}  \wedge \element{v}{\hat{d}} \, \} \)
\end{example}

\begin{example}   \label{Rhathat}
\hspace{-2mm}\(
\doublehat{R} =_{\mbox{\footnotesize{def.}}} 
\hspace{-2mm}\begin{array}[t]{l}
\{ \tuple{x,y}| \exists u_1 ... \exists u_n \exists v_1 ... \exists v_n. \\
\hspace{6mm} u_1\hat{R}v_1 \wedge ... \wedge u_n\hat{R}v_n \wedge\\
\hspace{6mm} x = u_1 \cdot ... \cdot u_n \hspace{1mm} \wedge \\
\hspace{6mm} y = v_1 \cdot ... \cdot v_n \hspace{2mm} \}
\end{array}
\)
\end{example}

@@

Universal box modalites are defined as duals, $\Box_a x = (\Diamond_a
x^c)^c$.  For instance, Aly is certain that the coin is heads if and
only if she does not consider it possible that it is not heads.

Summing up, given an Epik program with $n$ agents, we construct a concrete guarded
string model $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_1,...,\Diamond_n}$.  The
elements of $K$ are interpreted as propositions.  $0^c$ is the set of worlds, and it
may be countably infinite.  $\Diamond_i$ is an epistemic modality for the $i$th
agent.  Or referring to the Kripke relations $\doublehat{R}_i$, the construction
defines a multi-agent Kripke frame $\tuple{0^c,\doublehat{R}_1,...,\doublehat{R}_n}$
(usually a countable one) from an Epik specification. The frame consists of a set of
worlds, and an epistemic-alternative relation for each agent.  These models are used
as a target for natural-language interpretation in Section 5 and Section 6, where we
obtain semantic values such as \sem{Amy knows that it's heads and Bob knows that Amy
  knows whether it is heads or tails, and does not know that it's heads} as elements
of $K$. Concretely the proposositions are sets of guarded strings (usually countable
ones), construed as sets of worlds as they figure in possible worlds semantics for
natural language.

>>>>

These operators combine to form the algebraic structure Kleene Algebra
with Tests (\textsc{Kat}), which has the algebraic signature
$\tuple{K,+,\cdot,*,\bar{},0,1}$.  We interpret each term in as a set
of guarded strings as indicated in Figure~\ref{fig:interp}. A bare
event $e$ is interpreted as the set of guarded strings allowed by
$\varphi$ and $P$. The operation $p + q$ is set union. The constant
$0$ is the empty set.  The constant 1 is the set of all truth
assignments obeying the constraint on truth assignments,
i.e. $\{10,01\}$ in the coin example.  The operation $\cdot$ is fusion
product raised to sets: $x \cdot y$ is the set of all defined fusion
products of an element of $x$ with and element of $y$.  The operation
$*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$, where $x^i$ is the
$i$-times product of $x$ with itself and $x^0=1$.  Subsets of $1$
indicated by $t$ are also elements of $K$, and these form the Boolean
algebra of tests, denoting all atoms admitted by the constraint
$\varphi$ and the boolean test $t$.  We ``overload'' $\cdot$ to
further indicate boolean conjunction, $+$ to further indicate boolean
disjunction, and $0,1$ to indicate the zero and one elements of the
boolean algebra. The overbar operation denotes complement in this
Boolean algebra: $\overline t$ denotes all atoms $\alpha$ such that
$\alpha \Rightarrow \neg t$.  Other boolean operators (such as
implication $\Rightarrow$) are encoded as syntactic sugar in the
standard way.

To this we add an unary operation $\Diamond_a$ for each agent $a$, and a
unary complent operation $\cdot^{c}$ on terms.  This results in the
signature $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$
for epistemic KAT with two agents.  We interpret the epistemic
epistemic modality $\Diamond_a$ using Kripke semantics, as pre-image
relative to a fixed relation $R_a$ between guarded strings,
$\Diamond_a~x = \{u \mid \exists v. v \in x \wedge u \mathrel{R_a} v\}$.
Here $u$ and $v$ are guarded strings, while $x$ is an element of $K$.

@@

It remains to define the Kripke relation on guarded strings from an
agent specification as in Figure \ref{epikfigure}. In Epik, an agent
specification pairs each bare event with a set of ($+$-separated) bare
events, and so determines a relation between bare events, call it
relation $R_a$ for an agent $a$.  We can lift this to a relation
$\hat{R_a}$ between decorated events for a given $\varphi$ and $P$,
see \ex{Rhat}, and again to generalized to a relation $\doublehat{R}$
on arbitrary guarded strings by relational Kleene star (as opposed to
the Kleene star from aboce) , as defined in \ex{Rhathat}. The
operation $\cdot$ in the definition is KAT product, which enforces
matching of tests.

%% \begin{example}   \label{Rhat}
%% \hspace{-2mm} \(
%% \hat{R} \triangleq
%% \{\tuple{u,v} \mid \exists c \exists d. c\mathrel R d \wedge u \in \hat{c}  \wedge v \in \hat{d} \, \} \)
%% \end{example}

\begin{example}   \label{Rhathat}
\hspace{-2mm}\(
\doublehat{R} \triangleq
\hspace{-2mm}\begin{array}[t]{l}
\{ \tuple{x,y} \mid \exists u_1 ... \exists u_n \exists v_1 ... \exists v_n. \\
\hspace{6mm} u_1\mathrel{R}v_1 \wedge ... \wedge u_n\mathrel{R}v_n \wedge\\
\hspace{6mm} x \in \llbracket u_1 \cdot ... \cdot u_n \rrbracket^{\varphi,P}\hspace{1mm} \wedge \\
\hspace{6mm} y \in \llbracket v_1 \cdot ... \cdot v_n \rrbracket^{\varphi,P} \hspace{2mm} \}
\end{array}
\)
\end{example}

@@

The complement operation ${p}^c$ is complement at the level of sets of
guarded strings, defined to every set the universe $\_^*$ except those
denoted by $p$. Note that $\llbracket 0^c \rrbracket^{\varphi,P} =
\llbracket\_^*\rrbracket^{\varphi,P}$. We use the complement to define
universal box modalites as the dual of the diamond modality, $\Box_a x
= (\Diamond_a x^c)^c$.  For instance, Aly is certain that the coin is
heads if and only if she does not consider it possible that it is not
heads.

Summing up, given an Epik program with $n$ agents, we construct a
concrete guarded string model
$\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_{a_1},...,\Diamond_{a_n}}$.
The elements of $K$ are interpreted as propositions.  $0^c$ is the set
of worlds, and it may be countably infinite.  $\Diamond_{a_i}$ is an
epistemic modality for the $i$th agent.  Or referring to the Kripke
relations $\doublehat{R}_i$, the construction defines a multi-agent
Kripke frame $\tuple{0^c,\doublehat{R}_1,...,\doublehat{R}_n}$
(usually a countable one) from an Epik specification. The frame
consists of a set of worlds, and an epistemic-alternative relation for
each agent.  These models are used as a target for natural-language
interpretation in Section 5 and Section 6, where we obtain semantic
values such as \sem{Amy knows that it's heads and Bob knows that Amy
  knows whether it is heads or tails, and does not know that it's
  heads} as elements of $K$. Concretely the proposositions are sets of
guarded strings (usually countable ones), construed as sets of worlds
as they figure in possible worlds semantics for natural language.
